Community Detection Methods for Generalized Random Dot Product Graphs

### 2021-05-26 proposal/proposal.tex
updated slides

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Community Detection Methods for the Generalized Random Dot Product Graph Model}}\\
\vspace*{2\baselineskip}
\normalsize{A dissertation proposal submitted in partial satisfaction of the requirements for the degree of \\}
Doctor of Philosophy \\
in \\
Statistical Science \\
\vspace*{2\baselineskip}
\Large{John Koo}\\
\vspace*{3\baselineskip}
\Large{\textbf{Research Committee Members}}\\
Dr. Michael Trosset \\
Dr. Minh Tang \\
Dr. Julia Fukuyama \\
Dr. Roni Khardon \\
Dr. Fangzheng Xie \\
\vspace*{3\baselineskip}
Date TBA \\
\vspace*{1\baselineskip}
Department of Statistics \\
Indiana University \\
Bloomington, Indiana \\
\end{center}
% \end{titlepage}

@@

@@

\hypertarget{alternative-approaches-to-community-detection}{%
\subsection{Alternative Approaches to Community
Detection}\label{alternative-approaches-to-community-detection}}

>>>>

%\begin{titlepage}
\begin{center}
\LARGE{\textbf{Community Detection Methods for the Generalized Random Dot Product Graph Model}}\\
\vspace*{2\baselineskip}
\normalsize{A dissertation proposal submitted in partial satisfaction of the requirements for the degree of \\}
Doctor of Philosophy \\
in \\
Statistical Science \\
\vspace*{2\baselineskip}
\Large{John Koo}\\
\vspace*{3\baselineskip}
\Large{\textbf{Research Committee Members}}\\
Dr. Michael Trosset \\
Dr. Minh Tang \\
Dr. Julia Fukuyama \\
Dr. Roni Khardon \\
Dr. Fangzheng Xie \\
\vspace*{3\baselineskip}
June 2, 2021 \\
\vspace*{1\baselineskip}
Department of Statistics \\
Indiana University \\
Bloomington, Indiana \\
\end{center}
% \end{titlepage}

@@

\begin{center}
\LARGE{Abstract}
\end{center}

\vspace*{2\baselineskip}

\normalsize

Graph and network data, in which samples are represented not as a collection of feature vectors but as relationships between pairs of observations, are increasingly widespread in various fields ranging from sociology to computer vision. One common goal of analyzing graph data is community detection or graph clustering, in which the graph is partitioned into disconnected subgraphs in an unsupervised yet meaningful manner (e.g., by optimizing an objective function or recovering unobserved labels). Because traditional clustering techniques were developed for data that can be represented as vectors, they cannot be applied directly to graphs. In this research, we investigate the use of a family of spectral decomposition based approaches for community detection in block models, first by demonstrating how under the Generalized Random Dot Product Graph framework, all graphs generated by block models can be represented as feature vectors, then by proposing clustering methods for these feature vector representations and deriving the asymptotic properties of these methods.

@@

\hypertarget{non-spectral-approaches-to-community-detection}{%
\subsection{Non-Spectral Approaches to Community
Detection}\label{non-spectral-approaches-to-community-detection}}

### 2021-05-26 proposal/slides.Rmd
updated slides

---
title: Community Detection Methods for the Generalized Random Dot Product Graph Model
subtitle: Dissertation Proposal
author: John Koo
date: 'June 2021'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

@@

@@

1. Define a generative model 
$G \mid z_1, ..., z_n, \vec{\theta} \sim P(\vec{z}, \vec{\theta})$.
2. Develop a method for obtaining estimators $f(G) = \hat{z}_1, ..., \hat{z}_n$.
3. Identify asymptotic properties of estimators and prove consistency.

@@

1. Probability Models for Networks
    * Block Models and Community Structure
    * (Generalized) Random Dot Product Graphs
    * Connecting Block Models to the (G)RDPG

@@

2. Popularity Adjusted Block Model
    * Connecting the PABM to the GRDPG
    * Subspace Clustering for Community Detection
    * Orthogonal Spectral Clustering

@@

3. Community Detection for the (G)RDPG
    * Manifold Clustering
    * Manfolds as (G)RDPG Latent Configurations

@@

$A \sim BernoulliGraph(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \stackrel{ind}{\sim} Bernoulli(P_{ij})$ for each $i < j$.

Example: If $G$ is an Erdos-Renyi graph, then $P_{ij} = \theta$.

@@

@@

Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim BernoulliGraph(P)$ is a *block model*.

Example 1: Stochastic Block Model with two communities

@@

Random Dot Product Graph $A \sim RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $A \sim BernoulliGraph(X X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

Generalized Random Dot Product Graph $A \sim GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = blockdiag(I_p, -I_q)$
* $A \sim BernoulliGraph(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

If latent vectors $X_1, ..., X_n \stackrel{iid}{\sim} F$, then we write 
$(A, X) \sim RDPG(F, n)$ or $(A, X) \sim GRDPG_{p, q}(F, n)$.

@@

All Bernoulli Graphs are RDPG (if $P$ is 
positive semidefinite) or GRDPG.

Example 1 (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$

@@

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z))
```

@@

Example 1 (cont'd)

If we want to perform community detection,

1. Note that $A \sim BernoulliGraph(X X^\top)$ is a RDPG
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$  
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$

```{r, fig.height = 4, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

@@

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.

@@

**Notation**: $A \sim PABM(\{\lambda^{(kl)}\}_K)$.

@@

**Theorem** (KTT): 
Given popularity vectors $\{\lambda^{(kl)}\}_2$, we can define block diagonal $X \in \mathbb{R}^{n \times 4}$ and $U \in \mathbb{O}(4)$ such that $A \sim PABM(\lambda^{(kl)}_2)$ is equivalent to $A \sim GRDPG(X U)$.

@@

**Theorem** (KTT): $A \sim PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$
* $U \in \mathbb{O}(K^2)$
* $X \in \mathbb{R}^{n \times K^2}$ is block diagonal and 
composed of $\{\lambda^{(kl)}\}_K$ 
with each block corresponding to a community  
$\implies$ if $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $x_i^\top x_j = 0$.

**Remark** (non-uniqueness of the latent configuration):  
If $A \sim GRDPG_{p, q}(X)$, then $A \sim GRDPG_{p, q}(X Q)$ for any $Q$ in 
the indefinite orthogonal group with signature $p, q$.

@@

**Corollary**: Since $X$ is block diagonal by community and $U$ is orthogonal,
each community corresponds to a subspace in $\mathbb{R}^{K^2}$.

Subspace property holds even with linear transformation 
$Q \in \mathbb{O}(p, q)$.

Sparse Subspace Clustering algorithm:

1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
subject to $x_i = X c$ and $c_i^{(i)} = 0$.
2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
3. Construct affinity matrix $B = |C| + |C^\top|$.

Step (1) of SSC typically performed via LASSO: 
$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i} c\|_2^2 + \lambda \|c\|_1$

@@

If $X$ obeys the *Subspace Detection Property*, then $B$ is sparse such that 
$B_{ij} = 0$ if $x_i$ and $x_j$ belong to different subspaces

@@

qgraph::qgraph(sign(W), layout = 'spring', groups = factor(z))
```

@@

$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } 
\hat{x}_i = \hat{X} c \text{ and } c^{(i)} = 0$$
$$A \approx \hat{X} I_{p, q} \hat{X}^\top$$

@@

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
then $B_{ij} = 0$ $\forall i, j$ in different communities.

* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
$A_n \sim BernoulliGraph(P_n)$.
* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$, 
then $B_{ij} = 0$ $\forall i, j$ in different communities.

Orthogonal Spectral Clustering algorithm:

$$
\max_{i, j} \hat{B}^{(ij)} = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$

Let $(A, X) \sim RDPG(F, n)$ such that

1. Define functions $f_1, ..., f_K$ such that 
$f_k : [0, 1]^r \mapsto \mathbb{R}^d$ and $f_k(t) \neq f_l(t)$ 
$\forall k, l \in [K]$. 
2. Sample labels
$Z_1, ..., Z_n \stackrel{iid}{\sim} Categorical(\pi_1, ..., \pi_K)$.
3. Sample $T_1, ..., T_n \stackrel{iid}{\sim} D$ with support $[0, 1]^r$. 
4. Set latent positions $X_i = f_{Z_i}(T_i)$ and 
$X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}^\top$.
5. $A \sim BernoulliGraph(X X^\top)$

Athreya et al. and Rubin-Delanchy et al.: we can approximate properties of 
the latent configurations via ASE.

General community detection method:  
Given $A$, $K$, and $d$ (or $p$ and $q$),

  1. Use ASE to approximate the latent configuration.
  2. Use the appropriate clustering algorithm for the form of the latent
  configuration (manifolds).

**Example**: 
Let $U_1, ..., U_{n_1}, U_{n_1 + 1}, ..., U_{n} \stackrel{iid}{\sim} Uniform(0, \cos \frac{\pi}{2} a)$, $f_1(t) = (t, 0)$, and $f_2(t) = (t, a)$. $X_i = f_1(U_i)$ for $i \leq n_1$ and $X_j = f_2(U_j)$ for $n_1 + 1 \leq j \leq n$. If we observe $X_1, ..., X_{n_1}, X_{n_1+1}, ..., X_{n}$, what approach will allow us to group the observations by segment?

1. Single linkage clustering will produce perfect community detection.
2. Any $\epsilon$-neighborhood graph with $\epsilon \leq a$ will consist of 
at least $K$ disjoint subgraphs such that no subgraph contains members of two 
different communities.

**Example**: 
Starting with the parallel segments as before, suppose instead of observing $X_1, ..., X_n$, we have noisy observations $X_1 + \xi_1, ..., X_n + \xi_n$ such that 
$\max_i \|\xi_i\| = \xi \leq a / 3$.

Then $\forall a \in (0, 1)$, $\delta \in (0, 1)$, $K \geq 2$, $\xi \leq a / 3$, 
$\exists N(a, \delta, K, \xi) < \infty$ such that when $\min_k n_k \geq N$, 
with probability at least $1 - \delta$,

1. Single linkage will produce perfect community detection.
2. Any $\epsilon$-neighborhood graph will consist of at least $K$ sub-graphs with no subgraph containing vertices from multiple communities.

This also holds for noisy points sampled from one-dimensional manifolds such that the manifolds are distance at least $a$ apart.

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), layout = 'spring')
title('RDPG')
```

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, 1:2] %*% diag(sqrt(A.eigen$values[1:2]))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

1. Show that the ASE of a random graph generated by these latent vectors 
produces the correct conditions for sufficiently large $n$.
2. Extend results to non-uniform distributions.
3. Extend results to multidimensional manifolds.
4. Relax condition for the minimum distance between manifolds.
5. Explore more robust clustering techniques for these latent configurations.
6. Extend results to the GRDPG.

@@

* $Q^\top Q \neq I$
* If $A \sim GRDPG_{p, q}(X)$, then $A \sim GRDPG_{p, q}(X Q)$
* $(Q x)^\top (Q y) = x^\top Q^\top Q y \neq x^\top y$
* $\|Q\| \neq 1$ $\implies$ $\|Q x - Q y\| \neq \|x - y\|$

## Community Detection in Block Models

@@

* $Z_i \stackrel{iid}{\sim} Categorical(\pi_1, ..., \pi_K)$
* $\lambda_{ik} \stackrel{ind}{\sim} Beta(a_{ik}, b_{ik})$

>>>>

---
title: Community Detection Methods for Generalized Random Dot Product Graphs
subtitle: Dissertation Proposal
author: John Koo
date: 'June 2021'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

@@

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

```{r out.width = '50%', fig.height = 3, fig.width = 4}
Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
P <- Pz$P
z <- Pz$clustering
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
```

@@

1. Define a generative model for graph
$G \mid z_1, ..., z_n, \vec{\theta} \sim P(\vec{z}, \vec{\theta})$.
2. Develop a method for obtaining estimators $f(G) = \hat{z}_1, ..., \hat{z}_n$.
3. Identify asymptotic properties of estimators.

@@

1. Probability Models for Networks
    a. Block Models and Community Structure
    b. (Generalized) Random Dot Product Graphs
    c. Connecting Block Models to the (G)RDPG

@@

2. Popularity Adjusted Block Model
    a. Connecting the PABM to the GRDPG
    b. Subspace Clustering for Community Detection
    c. Orthogonal Spectral Clustering

@@

3. Community Detection for the (G)RDPG
    a. Manifold Clustering
    b. Manifolds as (G)RDPG Latent Configurations

@@

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

@@

**Example 1**: If every entry $P_{ij} = \theta$, then $A \sim \BG(P)$ is an Erdos-Renyi graph.

@@

Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim \BG(P)$ is a *block model*.

**Example 2**: Stochastic Block Model with two communities

@@

Random Dot Product Graph $A \sim \RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $A \sim \BG(X X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
* $A \sim \BG(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

If latent vectors $X_1, ..., X_n \iid F$, then we write 
$(A, X) \sim \RDPG(F, n)$ or $(A, X) \sim \GRDPG_{p, q}(F, n)$.

@@

All Bernoulli Graphs are RDPG (if $P$ is 
positive semidefinite)  
or GRDPG (in general).

**Example 2** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$

@@

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

@@

**Example 2** (cont'd): If we want to perform community detection,

1. Note that $A \sim \BG(P)$ is a RDPG since $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

@@

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.

@@

**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.

@@

**Theorem** (KTT): 
Given popularity vectors $\{\lambda^{(kl)}\}_2$, we can define block diagonal $X \in \mathbb{R}^{n \times 4}$ and $U \in \mathbb{O}(4)$ such that $A \sim \PABM(\{\lambda^{(kl)}\}_2)$ is equivalent to $A \sim \GRDPG_{3, 1}(X U)$.

@@

**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$
* $U \in \mathbb{O}(K^2)$
* $X \in \mathbb{R}^{n \times K^2}$ is block diagonal and 
composed of $\{\lambda^{(kl)}\}_K$ 
with each block corresponding to a community.

**Remark 1** (orthogonality of subspaces):
If $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $x_i^\top x_j = 0$.

**Remark 2** (non-uniqueness of the latent configuration):  
If $A \sim \GRDPG_{p, q}(Y)$, then $A \sim \GRDPG_{p, q}(Y Q)$ for any $Q$ in 
the indefinite orthogonal group with signature $p, q$.

**Remark 3**: Communities correspond to subspaces even with linear transformation $Q \in \mathbb{O}(p, q)$, but this may break the orthogonality property.

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
then $B_{ij} = 0$ if $z_i \neq z_j$.

**Algorithm**: Orthogonal Spectral Clustering:

$$
\max_{i, j} \hat{B}^{(ij)}_n = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$

@@

**Corollary**: The ASE of $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ lies near a collection of $K$-dimensional subspaces in $K^2$ dimensions.

**Algorithm**: Sparse Subspace Clustering:

1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
subject to $x_i = X c$ and $c^{(i)} = 0$.  
This is typically performed via LASSO: 
$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i} c\|_2^2 + \lambda \|c\|_1$
2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
3. Construct affinity matrix $B = |C| + |C^\top|$.

@@

Stronger result: Apply SSC to the eigenvectors of $A$.

$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } 
\hat{v}_i = \hat{V} c \text{ and } c^{(i)} = 0$$
$$A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$$

* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
$A_n \sim \BG(P_n)$.
* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.

## Simulation Study

```{r clust_err_k, fig.width = 10, fig.height = 3, out.width = '100%'}
setwd('~/dev/pabm-grdpg')
clustering.df <- readr::read_csv('clustering-k.csv')
ssc.df <- readr::read_csv('clustering-ssc-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error),
    first.q = quantile(error, .25),
    third.q = quantile(error, .75),
    med.err.ssc = median(error.ssc),
    first.q.ssc = quantile(error.ssc, .25),
    third.q.ssc = quantile(error.ssc, .75),
    med.err.ep = median(error.ep, na.rm = TRUE),
    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  scale_y_log10() +
  labs(y = 'community detection error rate', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC-ASE'), width = .1) + 
  # geom_line(aes(x = n, y = med.err.ep,
  #               colour = 'MM-EP')) + 
  # geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
  #                   colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
                    colour = 'SSC-A'), width = .1) + 
  geom_line(aes(x = n, y = med.err.louvain,
                colour = 'MM-Louvain')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
                    colour = 'MM-Louvain'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), layout = 'spring', legend = FALSE)
title('(G)RDPG')
```

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, c(1, 2)] %*% diag(sqrt(abs(A.eigen$values[c(1, 2)])))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

Let $(A, X) \sim \GRDPG_{p, q}(F, n)$ as follows:

1. Define functions $\gamma_1, ..., \gamma_K$ such that each 
$\gamma_k : [0, 1]^r \mapsto \mathbb{R}^d$ and $\gamma_k(t) \neq \gamma_l(t)$ 
when $k \neq l$.
2. Sample labels
$Z_1, ..., Z_n \iid \Cat(\pi_1, ..., \pi_K)$.
3. Sample $T_1, ..., T_n \iid D$ with support $[0, 1]^r$. 
4. Set latent positions $X_i = \gamma_{Z_i}(T_i)$ and 
$X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}^\top$.
5. $A \sim \BG(X I_{p, q} X^\top)$

Recall that the ASE approximates the true latent positions:

$$\max_i \|\hat{X}_i - Q_n X_i\| \stackrel{a.s.}{\to} 0$$

This leads to a general community detection method:

Given $A$, $K$, and $d$ (or $p$ and $q$),

  1. Use ASE to approximate the latent configuration.

  2. Use an appropriate clustering algorithm for the latent configuration.

**Example 3**: 
Let $U_1, ..., U_{n_1}, U_{n_1 + 1}, ..., U_{n} \iid \Uniform(0, \theta)$, $\gamma_1(t) = (t, 0)$, and $\gamma_2(t) = (t, a)$. $X_i = \gamma_1(U_i)$ for $i \leq n_1$ and $X_j = \gamma_2(U_j)$ for $n_1 + 1 \leq j \leq n$.  
If we observe $X_1, ..., X_{n_1}, X_{n_1+1}, ..., X_{n}$, what approach will allow us to group the observations by segment?

1. Single linkage clustering entails perfect community detection.

2. Any $\epsilon$-neighborhood graph with $\epsilon \leq a$ will consist of 
at least $K$ disjoint subgraphs such that no subgraph contains members of two 
different communities.

**Example 4**: 
Starting with the parallel segments as before, suppose instead of observing $X_1, ..., X_n$, we have noisy observations $X_1 + \xi_1, ..., X_n + \xi_n$ such that 
$\max_i \|\xi_i\| = \xi \leq a / 3$.

Then we can derive similar statements as in Example 3.

This also holds for noisy points sampled uniformly on one-dimensional manifolds such that the manifolds are distance at least $a$ apart.

Since the ASE of a RDPG generated from points on these segments/curves will reconstruct the original segments/curves with noise, we can apply this to the RDPG (if sufficient $n$).

1. Derive formal statements showing that the ASE of a random graph generated by these latent configurations produces the correct conditions for community detection.

2. Extend results to non-uniform distributions.

3. Extend results to multidimensional manifolds.

4. Relax condition for the minimum distance between manifolds.

5. Explore the performance of these techniques in the presence of outliers.

6. Extend results to the GRDPG.

## Subspace Detection Property

If $X$ obeys the *Subspace Detection Property*, then $B$ is sparse such that 
$B_{ij} = 0$ if $x_i$ and $x_j$ belong to different subspaces.

@@

qgraph::qgraph(sign(W), layout = 'spring', groups = factor(z), legend = FALSE)
```

@@

@@

@@

* $Q^\top Q \neq I$
* If $A \sim \GRDPG_{p, q}(X)$, then $A \sim \GRDPG_{p, q}(X Q)$
* $(Q x)^\top (Q y) = x^\top Q^\top Q y \neq x^\top y$
* $\|Q\| \neq 1$ $\implies$ $\|Q x - Q y\| \neq \|x - y\|$

## Non-Spectral Community Detection for Block Models

@@

* $Z_i \iid Categorical(\pi_1, ..., \pi_K)$
* $\lambda_{ik} \indep Beta(a_{ik}, b_{ik})$
